Map{R}educe in {MPI} for Large-scale Graph Algorithms

### 2010-11-04 paper/graph.tex
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@663 3fc257bb-8550-0410-81b3-a5c103da83f7

Note that the degree distribution of vertices in the graph depends on
the choice of parameters $a$, $b$, $c$, $d$.  If one of the four
values is larger than the other three, a skewed distribution results.
Variants of the above algorithm can be used when $N$ is not a
power-of-two, to generate graphs with weighted edges (assign a numeric
value to the $G_{ij}$ edge), graphs without self edges (require $i \ne
j$), or graphs with undirected edges (require $i < j$).  Or the
general R-MAT matrix can be further processed by MapReduce operations
to meet these requirements.  For some algorithms below (e.g., triangle
finding and maximal independent set generation), we post-process R-MAT
matrices to create upper-triangular R-MAT matrices representing
undirected graphs.  For each edge $(V_i,V_j)$ with $i > j$, we add
edge $(V_j,V_i)$, essentially symmetrizing the matrix; we then remove
duplicate edges and self edges, leaving only edges with $i < j$.

By using an additional MapReduce object, we can improve the
performance of the R-MAT generation algorithm, particularly when
out-of-core operations are needed.  In the enhanced algorithm
(Figure~\ref{fig:rmat2}), we emit newly created edges into MapReduce
object $G_{new}$ and {\it aggregate()} the edges to processors.  We
then add those edges to the MapReduce object $G$, containing all
previously generated unique edges; the edges of $G$ are already
aggregated to processors using the same mapping as $G_{new}$.  We can
then use the {\it compress()} operation to reduce duplicates locally
on each processor.  The cost of the enhanced algorithm is less due to
the smaller number of KV pairs communicated in the {\it aggregate()},
compared to the {\it collate()} operation of the original algorithm,
particularly in later iterations of the algorithm where few new edges
are generated.

>>>>

Note that the degree distribution of vertices in the graph depends on
the choice of parameters $a$, $b$, $c$, $d$.  If one of the four
values is larger than the other three, a skewed distribution results.
Variants of the algorithm can be used when $N$ is not a power-of-two,
to generate graphs with weighted edges (assign a numeric value to the
edge), graphs without self edges (require $i \ne j$), or graphs with
undirected edges (require $i < j$).  Or the general R-MAT matrix can
be further processed by MapReduce operations to meet these
requirements.  For some algorithms below (e.g., triangle finding and
maximal independent set generation), we post-process R-MAT matrices to
create upper-triangular R-MAT matrices representing undirected graphs.
For each edge $(V_i,V_j)$ with $i > j$, we add edge $(V_j,V_i)$,
essentially symmetrizing the matrix; we then remove duplicate edges
and self edges, leaving only edges with $i < j$.

By using an additional MapReduce object, we can improve the
performance of the R-MAT generation algorithm, particularly when
out-of-core operations are needed.  In the enhanced algorithm
(Figure~\ref{fig:rmat2}), we emit newly created edges into MapReduce
object $G_{new}$ and {\it aggregate()} the edges to processors.  We
then add those edges to the MapReduce object $G$, containing all
previously generated unique edges; the edges of $G$ are already
aggregated to processors using the same mapping as $G_{new}$.  We can
use the {\it compress()} operation to reduce duplicates locally on
each processor.  The enhanced algorithm runs faster due to the smaller
number of KV pairs communicated in the {\it aggregate()}, compared to
the {\it collate()} operation of the original algorithm, particularly
in later iterations of the algorithm where few new edges are
generated.
